The Forward March of Categorical Tolerance in the United States\thanks{Reserved for acknowledgments}

### 2025-06-08 manuscript.tex

\section*{Results}
\subsection*{Probability of Dislikes by Genre}
Table~\ref{tab:dis-by-genres} shows the probability of any genre experiencing a judgment of dislike. As we can see, Metal is the genre more likely to be disliked, consistent with previous work in the United States \citep{bryson1996anything-311, lizardo2015musical-8c6}. Electronic Dance Music, Rap/Hip Hop, and Broadway/Musicals are also toward the top in terms of dislikes. R\&B, Rock, and Pop/Top 40, on the other hand, have comparatively low odds of being disliked.

\subsection*{Univariate Distribution of Number of Dislikes}
\begin{figure}[ht!]
        \includegraphics[width=1.0\textwidth]{Tabs/desc-tab-dislike.png}
    \caption{Proportion of dislikes by genre.}
    \label{tab:dis-by-genres}
\end{figure}

\begin{figure}[ht!]
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol.png}
    \caption{Univariate plot of number of dislikes.}
    \label{fig:main}
\end{figure}
Figure~\ref{fig:main} shows the univariate distribution of the number of dislikes for the LM 2025 data. As we can see, the distribution exhibits excess zeroes, with about 20\% of the sample reporting zero dislikes among the twelve musical genres included.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-pol.png}
            \caption{Political Ideology}
            \label{fig:cat-tol-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-edu.png}
            \caption{Education}
            \label{fig:cat-tol-edu}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-inc.png}
            \caption{Income}
            \label{fig:cat-tol-inc}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-age.png}
            \caption{Age}
            \label{fig:cat-tol-age}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-rac.png}
            \caption{Ethnoracial Identification}
            \label{fig:cat-tol-rac}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-gen.png}
            \caption{Gender Identification}
            \label{fig:cat-tol-gen}
    \end{subfigure}
    \caption{Proportion of Categorical Tolerants across different socio-demographic characteristics.}
    \label{fig:cat-tol}
\end{figure}

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-pol.png}
            \caption{Political Ideology}
            \label{fig:grd-int-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-edu.png}
            \caption{Education}
            \label{fig:grd-int-edu}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-inc.png}
            \caption{Income}
            \label{fig:grd-int-inc}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-age.png}
            \caption{Age}
            \label{fig:grd-int-age}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-rac.png}
            \caption{Ethnoracial Identification}
            \label{fig:grd-int-rac}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-gen.png}
            \caption{Gender Identification}
            \label{fig:grd-int-gen}
    \end{subfigure}
    \caption{Average number of cultural dislikes across different socio-demographic characteristics.}
    \label{fig:grd-int}
\end{figure}
\newpage
\bibliography{references}
\bibliographystyle{apalike}

>>>>

\section*{Results}
\subsection*{Probability of Dislikes by Genre}
\begin{table}[ht!]
    \caption{Proportion of dislikes by genre.}
    \includegraphics[trim={0 11cm 0 0},clip, width=1.0\textwidth]{Tabs/desc-tab-dislike.png}
    \label{tab:dis-by-genres}
\end{table}

Table~\ref{tab:dis-by-genres} shows the probability of any genre experiencing a judgment of dislike. As we can see, Metal is the genre more likely to be disliked, consistent with previous work in the United States \citep{bryson1996anything-311, lizardo2015musical-8c6}. Electronic Dance Music, Rap/Hip Hop, and Broadway/Musicals are also toward the top in terms of dislikes. R\&B, Rock, and Pop/Top 40, on the other hand, have comparatively low odds of being disliked.

\subsection*{Univariate Distribution of Number of Dislikes}

\begin{figure}[ht!]
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol.png}
    \caption{Univariate plot of number of dislikes.}
    \label{fig:main}
\end{figure}
Figure~\ref{fig:main} shows the univariate distribution of the number of dislikes for the LM 2025 data. As we can see, the distribution exhibits excess zeroes, with about 20\% of the sample reporting zero dislikes among the twelve musical genres included. This figure is higher than the 15\% reported by \citet{lizardo2016end-4fb}. We should be cautious when making direct comparisons, given that the number of genres included is lower in the LM 2025 marketplace survey. Nevertheless, we can at the very least conclude that categorical tolerance remains a pervasive phenomenon among a substantial number of Americans. The taste regime governing patterns of cultural choice among Americans in the second decade of the twenty-first century certainly no longer resembles that captured by the GSS in the early 1990s, with a larger-than-expected number of people refusing to express dislikes for any genre.

\subsection*{Socio-Demographic Distribution of Categorical Tolerance}
\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-pol.png}
            \caption{Political Ideology}
            \label{fig:cat-tol-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-edu.png}
            \caption{Education}
            \label{fig:cat-tol-edu}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-inc.png}
            \caption{Income}
            \label{fig:cat-tol-inc}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-age.png}
            \caption{Age}
            \label{fig:cat-tol-age}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-rac.png}
            \caption{Ethnoracial Identification}
            \label{fig:cat-tol-rac}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-cat-tol-gen.png}
            \caption{Gender Identification}
            \label{fig:cat-tol-gen}
    \end{subfigure}
    \caption{Proportion of Categorical Tolerants across different socio-demographic characteristics.}
    \label{fig:cat-tol}
\end{figure}

Figure~\ref{fig:cat-tol} shows how the sample probability of being a categorical tolerant is distributed among various socio-demographic categories. As Figure~\ref{fig:cat-tol-age} shows, age is the most straightforward structuring factor, replicating the pattern observed by \citet{lizardo2016end-4fb}: Younger people are more likely to be categorical tolerants compared to older people. Nevertheless, the probability of being a categorical tolerant in these data is maximized among younger adults (individuals in their thirties), suggesting that younger people are induced to use musical genres for purposes of symbolic inclusion to a greater extent than people at this life-stage. Regardless, it is clear that if cohort replacement dynamics drive the rise of categorical tolerance, then we should expect it to remain a fixture of the cultural taste landscape in the U.S. as older, less categorically tolerant cohorts exit the population.

\subsection*{Socio-Demographic Distribution of Symbolic Exclusion}
\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-pol.png}
            \caption{Political Ideology}
            \label{fig:grd-int-pol}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-edu.png}
            \caption{Education}
            \label{fig:grd-int-edu}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-inc.png}
            \caption{Income}
            \label{fig:grd-int-inc}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-age.png}
            \caption{Age}
            \label{fig:grd-int-age}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-rac.png}
            \caption{Ethnoracial Identification}
            \label{fig:grd-int-rac}
    \end{subfigure}
     \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/uni-dist-grd-int-gen.png}
            \caption{Gender Identification}
            \label{fig:grd-int-gen}
    \end{subfigure}
    \caption{Average number of cultural dislikes across different socio-demographic characteristics.}
    \label{fig:grd-int}
\end{figure}
\newpage
\bibliography{references}
\bibliographystyle{apalike}
